---
title: "Predictors of Illness in Golden Retrievers"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: flatly
    source_code: embed
runtime: shiny
---

```{r global, include = FALSE}

#### NEED 4 MORE PREDICTORS A LA GUIDELINES, WILL BE BEHAVIOR ONCE CLEANED
# Downloading and cleaning data packages
library(here) # helps importing data on different devices
library(dplyr) # pipe commands, data cleaning
library(lubridate) # converting to date type variable
library(stringr) # str_ functions
library(forcats) # factors

# Modeling packages
library(tidymodels) # modelling
library(vip) # variable importance plots

# Showcasing data packages
library(ggplot2) # plots
library(flexdashboard)
library(shiny) # shiny dashboard
library(DT)

# Helper
library(shinyjs)
```


```{r downloading_data}
# Import Data 
# My data can be obtained by making an account on this website:
# https://datacommons.morrisanimalfoundation.org/
# The file names have not been changed, but I put all of the data into a 
# folder called "DO_NOT_UPLOAD_DATA", so delete that if you don't have your 
# data in a folder also called that

# Age
profile_data <- read.csv(here("DO_NOT_UPLOAD_DATA", 
                             "dog_profile.csv"))

# Activity, shade vs sun, etc., in a subfolder called "activity_lifestyle"
# get rid of "activity_lifestyle" if you don't have your data in a folder
# called that
activity_data <- read.csv(here("DO_NOT_UPLOAD_DATA", 
                              "activity_lifestyle", 
                              "activity_overview.csv"))

# Presence of illness
conditions_data <- read.csv(here("DO_NOT_UPLOAD_DATA", 
                               "conditions_summary.csv"))

# Has energy score, excitability, trainability, dog rivalry, chasing
# in a subfolder called "behavior" get rid of "behavior" if you don't have your 
# data in a folder called that
behavior_data <- read.csv(here("DO_NOT_UPLOAD_DATA", 
                              "behavior", 
                              "behavior_summary.csv"))

# Home environment
environment_data <- read.csv(here("DO_NOT_UPLOAD_DATA", 
                                  "environment",
                                  "environment_conditions.csv"))


# Converting all dataframes to tibbles
profile_tbl <- as_tibble(profile_data)
activity_tbl <- as_tibble(activity_data)
conditions_tbl <- as_tibble(conditions_data)
behavior_tbl <- as_tibble(behavior_data)
environment_tbl <- as_tibble(environment_data)
```

```{r cleaning_profile_data}
# Making a variable called "starting age" using birth date and enrolled date
# lubridating all dates
profile_w_age <- profile_tbl %>%
  mutate(
    birth_date_new = ym(birth_date),
    enrolled_date_new = ym(enrolled_date),
    starting_age = as.integer(enrolled_date_new - birth_date_new)) %>%
  mutate(neuter_spay = str_trim(as.character(str_extract_all(sex_status, "(.*) "))),
         sex = str_trim(as.character(str_extract_all(sex_status, " (.*)"))),
         neuter_spay_fact = as_factor(case_when(neuter_spay == "Intact" ~ 0,
                                      neuter_spay != "Intact" ~ 1))) %>% 
  select(subject_id, # selecting only columns I am going to use
         neuter_spay_fact,
         sex,
         starting_age,
         birth_date_new,
         enrolled_date_new)

```

```{r cleaning_activity_data}
# Making characters into factors that are ordered specifically
activity_reordered <- activity_tbl %>%
  mutate(
    activity_level_reordered = fct(
      activity_level,
      c("",
        "None",
        "Little",
        "Moderate",
        "Very active")
    ),
    sun_exposure_duration_reordered = fct(
      sun_exposure_duration,
      c(
        "",
        "Less than 3 hours",
        "Between 3-8 hours",
        "Between 8-16 hours",
        "Between 9-16 hours",
        "More than 16 hours"
      )
    ),
    sun_exposure_shade_access_reordered = fct(
      sun_exposure_shade_access,
      c("",
        "No access to full shade",
        "Access to full shade")
    )
  ) %>%
  select( # selecting only columns I need
    subject_id,
    year_in_study,
    activity_level_reordered,
    sun_exposure_duration_reordered,
    sun_exposure_shade_access_reordered
  )

```

```{r cleaning_illness_data}
# Just want presence of illness of any given year
conditions <- conditions_tbl %>%
  filter(to_date == 0) %>%  # to date == 0 means that we are not including presence of sickness in previous years
  select(subject_id,
         year_in_study,
         any) %>%
  drop_na(any) %>% # need NO NAs for the model to work
  mutate(any = as_factor(any))
```


```{r cleaning_behavioral_data}
# Behavioral indicators (scored out of 4, 0 means not a lot, 4 means severe/ a lot)

# separation
behavior_separation <- behavior_tbl %>%
  filter(to_date == 0,
         topic == "score_separation_related_problems") %>%
  mutate(
    count_0_sep = count_0 / 8,
    count_1_sep = count_1 / 8,
    count_2_sep = count_2 / 8,
    count_3_sep = count_3 / 8,
    count_4_sep = count_4 / 8,
    count_na_sep = count_na / 8
  ) %>%
  select(
    subject_id,
    year_in_study,
    count_0_sep,
    count_1_sep,
    count_2_sep,
    count_3_sep,
    count_4_sep
  )

behavior_energy <- behavior_tbl %>%
  filter(to_date == 0,
         topic == "score_energy") %>%
  mutate(count_0_energy = count_0 / 2,
    count_1_energy = count_1 / 2,
    count_2_energy = count_2 / 2,
    count_3_energy = count_3 / 2,
    count_4_energy = count_4 / 2,
    count_na_energy = count_na / 2) %>%
  select(subject_id,
    year_in_study,
    count_0_energy,
    count_1_energy,
    count_2_energy,
    count_3_energy,
    count_4_energy)

```

```{r cleaning_environment_data}
# Taking the variables I want to use (binary is there smoke, is there aerosols, has the dog been to the ER)
environment_selected <- environment_tbl %>% 
  filter(to_date == 0) %>% 
  mutate(smoke_exposure_fact = as_factor(smoke_exposure),
         poison_clinic_visit_fact = as_factor(poison_clinic_visit),
         use_aerosol_fact = as_factor(use_aerosol)) %>% 
  select(subject_id,
         year_in_study,
         smoke_exposure_fact,
         poison_clinic_visit_fact,
         use_aerosol_fact)
```

```{r joining_data}
# Joining the data
data_y_all <- full_join(profile_w_age,
                        activity_reordered,
                        by = "subject_id") %>%
  full_join(conditions,
            by = c("subject_id",
                   "year_in_study")) %>%
  full_join(behavior_separation,
            by = c("subject_id",
                   "year_in_study")) %>%
  full_join(behavior_energy,
            by = c("subject_id",
                   "year_in_study")) %>% 
  full_join(environment_selected,
            by = c("subject_id",
                   "year_in_study")) %>% 
  na.omit()

```

Sidebar {.sidebar}
===========================================================

<br>

`r h4("Objective:")`
`r h4("Predict whether a golden retriever will have an illness present based on selected inputs.")`

<br>

- - -

```{r function}
# adds all behavioral scores related to a topic to a list
topic_add <- function (input, topic) {
  if (str_detect(str_flatten(input, collapse = ", "),
                 str_c("count_0", topic, sep = "_"))) {
    input.new <- c(
      input,
      str_c("count_1", topic, sep = "_"),
      str_c("count_2", topic, sep = "_"),
      str_c("count_3", topic, sep = "_"),
      str_c("count_4", topic, sep = "_")
    )
    return(input.new)
  }
  else{
    return(input)
  }
}

```


```{r inputs}
# Year in Study as a reactive input
sliderInput(
  inputId = "year_input",
  label   = h5("Year in Study"),
  value   = min(data_y_all$year_in_study),
  min     = min(data_y_all$year_in_study),
  max     = max(data_y_all$year_in_study),
  step    = 1,
  ticks   = FALSE
)

br()

# Letting a use choose which variables to include in the model
checkboxGroupInput(
  inputId = "model_variables",
  label = h5("Variables to Include in Model"),
  choices = c(
    "sex" = "sex",
    "neutered/spayed, or intact?" = "neuter_spay_fact",
    "age at start of study" = "starting_age",
    "activity level" = "activity_level_reordered",
    "sun exposure" = "sun_exposure_duration_reordered",
    "does the dog have access to shade?" = "sun_exposure_shade_access_reordered",
    "does the dog have separation anxiety?" = "count_0_sep",
    "does the dog have high energy; do they act 'puppyish'?" = "count_0_energy",
    "does the dog get exposed to smoke?" = "smoke_exposure_fact",
    "has ingested poison or hazardous material requiring vet or emergency clinic visit?" = "poison_clinic_visit_fact",
    "exposed to aerosols in the home" = "use_aerosol_fact"),
  inline = FALSE
)




br()

actionButton(inputId = "apply", label = "Apply Changes", icon = icon("play"))
actionButton(inputId = "reset", label = "Reset", icon = icon("sync"))

br()
br()
br()

observeEvent(eventExpr = input$reset, handlerExpr = {
  
  updateSliderInput(
    session = session,
    inputId = "year_input",
    value   = min(data_y_all$year_in_study))
  
  #update if you figure out scoring!
  updateCheckboxGroupInput(
    session = session,
    inputId = "model_variables",
    selected = character(0))
  
  # makes reset complete
  shinyjs::delay(ms = 300, expr = {
    shinyjs::click(id ="apply")
  })
  
})

renderPrint(input$year_input)
renderPrint(input$model_variables)
```


```{r data_filtered}
# creating function golden_filtered(), which is reactive and lets the user change
# what year's data they are looking at

golden_filtered <- eventReactive(
  eventExpr = input$apply,
  valueExpr = {
    data_y_all %>%
      select(any,
             year_in_study,
             topic_add(topic_add(input$model_variables, "sep"), "energy")) %>%
      filter(year_in_study == input$year_input) # getting the datasets from specific years
  },
  ignoreNULL = FALSE
)
```

```{r data_splitting}
# Data splitting and re-sampling
set.seed(123)

splits <- eventReactive(
  eventExpr = input$apply,
  valueExpr = {
    initial_split(golden_filtered(), strata = any)
  },
  ignoreNULL = FALSE
)


golden_other <- reactive({
  training(splits())
})


golden_test  <- reactive({
  testing(splits())
})

# Create a validation set
set.seed(234)
prop.validation <- .10

val_set <- reactive({
  validation_split(golden_other(),
                   strata = any,
                   prop   = 1 - prop.validation)
})

```

```{r model}
lr_model_golden <- logistic_reg(mixture = 1, 
                                penalty = tune()) %>% 
  set_engine("glmnet")
```

```{r recipe}
lr_recipe_golden <- eventReactive(eventExpr = input$apply,
                        valueExpr = {
  recipe(any ~ ., data = golden_other()) %>%  
    step_dummy(all_nominal(), -all_outcomes()) %>% 
    step_zv(all_predictors()) %>% 
    step_normalize(all_predictors())
                          },
  ignoreNULL = FALSE
  )
```

```{r workflow}
lr_workflow_golden <- reactive({
  workflow() %>% 
  add_model(lr_model_golden) %>% 
  add_recipe(lr_recipe_golden())
})
```

```{r tuning_grid}
# `length.out` will deterrmine how many penalty values are tested in `tune_grid()`
lr_reg_grid_golden <- tibble(penalty = 10^seq(-4, -1, length.out = 10)) 


# Train and tune the model
lr_tune_golden <- reactive({
  lr_workflow_golden() %>% 
  tune_grid(resamples = val_set(), 
            grid      = lr_reg_grid_golden,
            control   = control_grid(save_pred = TRUE), # needed to get data for ROC curve
            metrics   = metric_set(roc_auc))
})


lr_best_golden <- reactive({
  lr_tune_golden() %>% 
  select_best(metric = "roc_auc")
})

```

```{r best_workflow}
# Add best validation model to workflow
lr_workflow_best_golden <- reactive({
  finalize_workflow(
  lr_workflow_golden(),
  lr_best_golden())       # needs to have the same column name as tune()
})

```

```{r fit_training}
# Inspect fit on entire training data
lr_fit_golden <- reactive({
  lr_workflow_best_golden() %>% 
  fit(golden_other())
})

```

```{r last_fit}
lr_last_fit_golden <- reactive({
  last_fit(
    lr_workflow_best_golden(),
    splits())
})

```

```{r confusion_matrix}
lr_conf_mat_golden <- reactive({
  lr_last_fit_golden() %>%
  collect_predictions() %>% 
  conf_mat(truth = any, estimate = .pred_class) 
})


```


Validation Summary
===========================================================


## Column {data-width="500"}

### Data Splitting

**Total Observations:**  
`r reactive(dim(golden_filtered())[1] %>% scales::comma())`

**Training Set:**  
`r reactive(dim(golden_other())[1] %>% scales::comma())`

**Validation Set:**  
`r reactive((dim(golden_other())[1] * prop.validation) %>% scales::comma())`

**Testing Set:**  
`r reactive(dim(golden_test())[1] %>% scales::comma())`


### Data Viewer

```{r table}
output$data_viewer <- DT::renderDataTable({
golden_filtered() %>% 
    slice(1:100) %>% 
    datatable(options = list(searching = FALSE,
                           pageLength = 50,
                           lengthMenu = c(50, 100),
                           scrollY = '250px',
                           scrollX = '300px'),
            style = "default")
})

DT::dataTableOutput(outputId = "data_viewer")
```


Column {data-width=650}
-----------------------------------------------------------------------

### Case Imbalance Check

```{r, eval = TRUE}
output$case_plot <- renderPlot({
  lr_recipe_golden() %>%
  prep() %>%
  juice() %>%
  ggplot(aes(any)) + 
    geom_bar(fill = 'darkgreen')
})

plotOutput(outputId = "case_plot")
```


Classification Results
===========================================================

### Confusion Matrix

```{r confusion}
output$conf_mat_plot_golden <- renderPlot({
  if (length(golden_filtered()) < 3) {
    plot(1,1,col="white")
    text(1,1,"You need to select variables for this plot to appear")
  } else{
  lr_conf_mat_golden() %>%
    autoplot(type = "heatmap")
  }

})

plotOutput(outputId = "conf_mat_plot_golden")

# Helped:
# https://stackoverflow.com/questions/30047010/shiny-custom-warning-error-messages 
```


### ROC AUC

```{r rocauc}
output$lr_auc_golden <- renderPlot({
   if (length(golden_filtered()) < 3) {
    plot(1,1,col="white")
    text(1,1,"You need to select variables for this plot to appear")
  } else{
  lr_tune_golden() %>%
  collect_predictions(parameters = lr_best_golden()) %>%
  roc_curve(any, .pred_0) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_path(color = 'green') +
  geom_abline(lty = 3, color = 'darkgreen') +
  coord_equal() +
  theme_classic()
  }
})

plotOutput(outputId = "lr_auc_golden")


renderPrint(lr_tune_golden()
            %>% collect_predictions(parameters = lr_best_golden()) )
```

Predictor Metrics
===========================================================

Column {data-width=350}
-----------------------------------------------------------------------

### Variable Importance Plot

```{r vip_plot}
output$vip_plot <- renderPlot({ 
   if (length(golden_filtered()) < 3) {
    plot(1,1,col="white")
    text(1,1,"You need to select variables for this plot to appear")
  } else {
  lr_fit_golden() %>% 
  pull_workflow_fit() %>%
  vip(num_features = 8) +
    theme_light()
  }
})

plotOutput(outputId = "vip_plot")
```

