---
title: "Predictors of Illness in Golden Retrievers"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: yeti
    source_code: embed
runtime: shiny
---

```{r global, include = FALSE}

#### NEED 4 MORE PREDICTORS A LA GUIDELINES, WILL BE BEHAVIOR ONCE CLEANED
library(here) # helps importing data on different devices
library(dplyr) # pipe commands, data cleaning
library(ggplot2) # plots
library(lubridate) # converting to date type variable
library(tidymodels) # modelling
library(janitor)
library(flexdashboard)
library(shiny) # shiny dashboard
library(vip) # variable importance plots
library(forcats) # factors
library(DT)

# Import Data 
# My data can be obtained by making an account on this website:
# https://datacommons.morrisanimalfoundation.org/
# The file names have not been changed, but I put all of the data into a 
# folder called "DO_NOT_UPLOAD_DATA", so delete that if you don't have your 
# data in a folder also called that

# Age
profile_data <- read.csv(here("DO_NOT_UPLOAD_DATA", 
                             "dog_profile.csv"))

# Activity, shade vs sun, etc., in a subfolder called "activity_lifestyle"
# get rid of "activity_lifestyle" if you don't have your data in a folder
# called that
activity_data <- read.csv(here("DO_NOT_UPLOAD_DATA", 
                              "activity_lifestyle", 
                              "activity_overview.csv"))

# Has energy score, excitability, trainability, dog rivalry, chasing
# in a subfolder called "behavior" get rid of "behavior" if you don't have your 
# data in a folder called that
behavior_data <- read.csv(here("DO_NOT_UPLOAD_DATA", 
                              "behavior", 
                              "behavior_summary.csv"))

conditions_data <- read.csv(here("DO_NOT_UPLOAD_DATA", 
                               "conditions_summary.csv"))



profile_tbl <- as_tibble(profile_data)
activity_tbl <- as_tibble(activity_data)
behavior_tbl <- as_tibble(behavior_data)
conditions_tbl <- as_tibble(conditions_data)

```


Sidebar {.sidebar}
===========================================================

<br>

`r h3("Objective:")`
`r h3("Predict whether a golden retriever will have an illness present based on selected inputs.")`

<br>

- - -

<br>

```{r cleaning}

# Making a variable called "starting age" using birth date and enrolled date
# lubridating all dates
profile_w_age <- profile_tbl %>% 
  mutate(birth_date_new = ym(birth_date),
         enrolled_date_new = ym(enrolled_date),
         starting_age = as.integer(enrolled_date_new - birth_date_new)) %>% 
  select(subject_id, sex_status, starting_age, birth_date_new, enrolled_date_new)

# Making characters into factors that are ordered specifically 
activity_reordered <- activity_tbl %>% 
  mutate(activity_level_reordered = fct(activity_level, 
                                  c("",
                                    "None",
                                    "Little",
                                    "Moderate",
                                    "Very active")),
         sun_exposure_duration_reordered = fct(sun_exposure_duration,
                                               c("",
                                                 "Less than 3 hours",
                                                 "Between 3-8 hours",
                                                 "Between 8-16 hours",
                                                 "Between 9-16 hours",
                                                 "More than 16 hours")),
         sun_exposure_shade_access_reordered = fct(sun_exposure_shade_access,
                                                   c("",
                                                     "No access to full shade",
                                                     "Access to full shade"))
         ) %>% 
  select(subject_id,
         year_in_study,
         activity_level_reordered, 
         sun_exposure_duration_reordered,
         sun_exposure_shade_access_reordered)


# Just want presence of illness of any given year
conditions <- conditions_tbl %>% 
  filter(to_date == 0) %>%  # to date == 0 means that we are not including presence of sickness in previous years
  select(subject_id,
         year_in_study,
         any) %>% 
  drop_na(any) %>%
  mutate(any = as_factor(any))


# Behavioral indicators (scored out of 4, 0 means not a lot, 4 means severe/ a lot)

# separation
behavior_separation <- behavior_tbl %>% 
  filter(to_date == 0,
         topic == "score_separation_related_problems") %>% 
  select(subject_id, 
         year_in_study, 
         count_0, 
         count_1, 
         count_2, 
         count_3, 
         count_4, 
         count_na)

data_y_all <- full_join(profile_w_age, 
                    activity_reordered, 
                    by = "subject_id") %>% 
  full_join(conditions,
            by = c("subject_id", 
                   "year_in_study")) %>% 
   full_join(behavior_separation,
            by = c("subject_id", 
                   "year_in_study"))
  na.omit()


```

```{r year_input}
# Year in Study as a reactive input
sliderInput(
  inputId = "year_input", 
  label   = h4("Year in Study"),
  value   = min(data_y_all$year_in_study),
  min     = min(data_y_all$year_in_study), 
  max     = max(data_y_all$year_in_study),
  step    = 1,
  ticks   = FALSE)

renderPrint(input$year_input)
```

```{r column_input}
checkboxGroupInput(
  inputId = "model_variables",
  label = h4("Variables to Include in Model"),
  choices = c("sex (including spay or neuter status)" = "sex_status",
              "age at start of study" = "starting_age",
              "activity level" = "activity_level_reordered",
              "sun exposure" = "sun_exposure_duration_reordered",
              "does the dog have access to shade?" = "sun_exposure_shade_access_reordered"),
  inline = FALSE
)

renderPrint(input$model_variables)
```


```{r data_filtered}
# creating function golden_filtered(), which is reactive and lets the user change
# what year's data they are looking at

golden_filtered <- reactive({
  data_y_all %>%
    select(any, year_in_study, input$model_variables) %>% 
    filter(year_in_study == input$year_input) # getting the datasets from specific years
})

renderPrint(dim(golden_filtered()))

```

```{r data_splitting}
# Data splitting and re-sampling
set.seed(123)

splits <- reactive({
  initial_split(golden_filtered(), strata = any)
})


golden_other <- reactive({
  training(splits())
})


golden_test  <- reactive({
  testing(splits())
})

# Create a validation set
set.seed(234)
prop.validation <- .20

val_set <- reactive({
  validation_split(golden_other(), 
                   strata = any, 
                   prop   = 1-prop.validation)
})

renderPrint(val_set())
```

```{r model}
lr_model_golden <- logistic_reg(mixture = 1, 
                                penalty = tune()) %>% 
  set_engine("glmnet")
```

```{r recipe}
lr_recipe_golden <- reactive({
  recipe(any ~ ., data = golden_other()) %>%  
    step_dummy(all_nominal(), -all_outcomes()) %>% 
    step_zv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors())
})
```

```{r workflow}
lr_workflow_golden <- reactive({
  workflow() %>% 
  add_model(lr_model_golden) %>% 
  add_recipe(lr_recipe_golden())
})
```

```{r tuning_grid}
# `length.out` will deterrmine how many penalty values are tested in `tune_grid()`
lr_reg_grid_golden <- tibble(penalty = 10^seq(-4, -1, length.out = 10)) 


# Train and tune the model
lr_tune_golden <- reactive({
  lr_workflow_golden() %>% 
  tune_grid(resamples = val_set(), 
            grid      = lr_reg_grid_golden,
            control   = control_grid(save_pred = TRUE), # needed to get data for ROC curve
            metrics   = metric_set(roc_auc))
})


lr_best_golden <- reactive({
  lr_tune_golden() %>% 
  select_best("roc_auc")
})

```

```{r best_workflow}
# Add best validation model to workflow
lr_workflow_best_golden <- reactive({
  finalize_workflow(
  lr_workflow_golden(),
  lr_best_golden())       # needs to have the same column name as tune()
})

```

```{r fit_training}
# Inspect fit on entire training data
lr_fit_golden <- reactive({
  lr_workflow_best_golden() %>% 
  fit(golden_other())
})

```

```{r last_fit}
lr_last_fit_golden <- reactive({
  last_fit(
    lr_workflow_best_golden(),
    splits())
})

```

```{r confusion_matrix}
lr_conf_mat_golden <- reactive({
  lr_last_fit_golden() %>%
  collect_predictions() %>% 
  conf_mat(truth = any, estimate = .pred_class) 
})


```


Validation Summary
===========================================================


## Column {data-width="500"}

### Data Splitting

**Total Observations:**  
`r reactive(dim(golden_filtered())[1] %>% scales::comma())`

**Training Set:**  
`r reactive(dim(golden_other())[1] %>% scales::comma())`

**Validation Set:**  
`r reactive((dim(golden_other())[1] * prop.validation) %>% scales::comma())`

**Testing Set:**  
`r reactive(dim(golden_test())[1] %>% scales::comma())`

### Data Viewer

```{r table}
output$data_viewer <- DT::renderDataTable({
golden_filtered() %>% 
    slice(1:100) %>% 
    datatable(options = list(searching = FALSE,
                           pageLength = 50,
                           lengthMenu = c(50, 100),
                           scrollY = '250px',
                           scrollX = '300px'),
            style = "default")
})

DT::dataTableOutput(outputId = "data_viewer")
```

Column {data-width=650}
-----------------------------------------------------------------------

### Model

```{r, eval = TRUE}
output$case_plot <- renderPlot({
  lr_recipe_golden() %>%
  prep() %>%
  juice() %>%
  ggplot(aes(any)) + # run into errors with starting age
    geom_bar(fill = 'darkgreen')
})

plotOutput(outputId = "case_plot")
```

```{r prediction}
output$lr_auc_golden <- renderPlot({
  lr_tune_golden() %>%
  collect_predictions(parameters = lr_best_golden()) %>%
  roc_curve(any, .pred_any) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_path(color = 'green') +
  geom_abline(lty = 3, color = 'darkgreen') +
  coord_equal() +
  theme_classic()
})

plotOutput(outputId = "lr_auc_golden")


renderPrint(lr_tune_golden() %>% collect_predictions(parameters = lr_best_golden()) )
```

Classification Results
===========================================================

### ROC AUC - makes the rest of stuff not show up

```{r rocauc}

lr_tune_golden() %>%
   collect_predictions(parameters = lr_best_golden())
  
# output$lr_auc_golden <- renderPlot({
#   lr_tune_golden() %>%
#   collect_predictions(parameters = lr_best_golden()) %>%
#   roc_curve(any, .pred_any) %>%
#   ggplot(aes(x = 1 - specificity, y = sensitivity)) +
#   geom_path(color = 'forestgreen') +
#   geom_abline(lty = 3, color = 'lightgreen') +
#   coord_equal() +
#   theme_classic()
# })
# 
# plotOutput(outputId = "lr_auc_golden")
# 
# renderPrint(lr_tune_golden() %>%
#               collect_predictions(parameters = lr_best_golden()) )
```
